---
title: "Computational Musicology"
author: Elias Groot
date: "still in progress"
output: 
  flexdashboard::flex_dashboard:
    theme: readable
    vertical_layout: scroll
---

```{r setup, include=FALSE}
library(flexdashboard)
library(knitr)
library(tidyverse)
library(spotifyr)
library(dplyr)
library(ggplot2)
remotes::install_github('jaburgoyne/compmus')
library(compmus)
```
Introduction
==============================
Column  
-----------------------------------------------------------------------

#### Hiphop now and then
In this course I want to do research on the difference and similarities between 'old school hiphop' from the 90s and the modern day rap and hiphop from around 2015 to 2023. I'm interested in how the dynamics have changed and if modern day rap for example contains more energy and danceability and less speechiness or acousticness compared to old school. Therefore my corpus will contain a merged group of tracks of both these periods of hiphop/rap. I will pick some modern day 'rap hits' and some very well known 'old school rap tracks'. 

The natural groups and comparison points will be these popular tracks from both these time spans and albums. I expect to see an increase in aspects like loudness because of the way music and hip hop beats get produced these days but I am very unsure how other factors like energy, valence and danceability have changed between these two groups. Besides things like this I am interested in how it differs in general or maybe how they have similarities because they both were popular in their own time. 

The strength of my corpus will be that I want to respectively check, with the help of popularity charts and checking the streams of the tracks on Spotify, which tracks were the most popular in these genres in both the 90s and around 2015 to 2023 and use these as a reference point. Hereby I rule out a certain form of subjectivity and my own taste of tracks because I will only focus on tracks that are popular and have a lot of streams. So despite that these two time spans differ a lot; my comparison groups have one thing in common, they both consist of only really popular tracks. 

```{r}

```



 

Visualisation    
================================
Column 
-----------------------------------------------------------------------

### Looking in to the SpotifyAPI track features. 
To begin the visualisation of our selected corpus we take a look at the correlation between energy and valence and how the differ from the 90s to the modern day rap music. Valence and energy are both variables that measure from 0.0 to 1.0. Valence referring to the emotional quality the music has: 1 being very uplifting and happy and 0.0 being angry, regretful or sad. Energy gives us an idea about the intensity and the activity of the track. A really energetic track would feel rousing for example.  

```{r, fig.width=8, fig.height=6}
library(tidyverse)
library(spotifyr)
library(plotly)
corpus <- get_playlist_audio_features("", "4QfiuYUQDJExOBuLtIVLuO")


library(plotly)

gg <- ggplot(corpus, aes(x = as.Date(track.album.release_date), y = valence, color = energy, text = paste(track.name))) +
    geom_point(size = 2) +
    scale_x_date(date_breaks = "years", date_labels = "%Y", expand = c(0, 0)) +
    xlim(as.Date("1990-01-01"), as.Date("2023-12-31")) + 
    scale_y_continuous(expand = c(0, 0), limits = c(0, 1)) +
    labs(x = "Release Date", y = "Valence",
         title = "Valence and Energy of Popular Hip Hop and Rap Songs", subtitle = "90s vs 2015-2023", color = "Energy") +
    scale_color_gradient(low = "blue", high = "red", limits = c(0,1)) +         theme_bw() +theme(panel.grid = element_blank(),
          plot.title = element_text(size = 18, face = "bold"),
          plot.subtitle = element_text(size = 14, face = "italic"),
          axis.title = element_text(size = 12),
          axis.text = element_text(size = 10),
          legend.title = element_text(size = 12),
          legend.text = element_text(size = 10))
    
ggplotly(gg)
```
***
Hoover on the points in the scatterplot to see the track names, exact release date and exact values for the variables energy and valence

### Chromagram of the outliers 
We see that there are some outliers here who are remarkable; let us take a look at the song "rockstar" from Post Malone featuring 21Savage. 
```{r, fig.width=8, fig.height=6}
library(tidyverse)
library(spotifyr)
library(compmus)
rockstar <-
  get_tidy_audio_analysis("0e7ipj03S05BNilyu5bRzt") |>
  select(segments) |>
  unnest(segments) |>
  select(start, duration, pitches)
rockstar |>
  mutate(pitches = map(pitches, compmus_normalise, "euclidean")) |>
  compmus_gather_chroma() |> 
  ggplot(
    aes(
      x = start + duration / 2,
      width = duration,
      y = pitch_class,
      fill = value
    )
  ) +
  geom_tile() +
  labs(x = "Time (s)", y = NULL, fill = "Magnitude") +
  theme_minimal() +
  scale_fill_viridis_c() +
  ggtitle("Chromagram from the song: rockstar (feat. 21 Savage)")
```

Conclusion
=============================
Column
-----------------------------------------------------------------------
### Chart C

```{r}

```

